---
title: "point_wise_errors"
output: html_document
---
# Setup
```{r setup}
setwd(rprojroot::find_rstudio_root_file())
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_current$set(root.dir = rprojroot::find_rstudio_root_file())

library(abind)
library(Rssa)
library(svd)
library(scales)
library(rTensor)
library(ggplot2)
library(dplyr)
library(tidyr)
library(zoo)
library(quantmod)
library(Rcpp)
library(RcppArmadillo)
library(future.apply)
library(progressr)

source(file = "./Source/TSSA.R")
source(file = "./Source/tsgen.R")
source(file = "./Source/utils.R")
```

```{r enable progressr, eval = FALSE}
# This commands should be executed in R console, not as a chunk
# Otherwise they fail
progressr::handlers(global = TRUE)
progressr::handlers("progress")
```

```{r}
pred_ssa_rec <- function() {
  if (is.complex(signal)) {
    noise <- CCSWGN(N, 0, sd)
  } else {
    noise <- rnorm(N, 0, sd)
  }
  x <- signal + noise
  rec <- reconstruct(ssa(x, L = L_ssa), groups = groups_ssa)
  Reduce("+", rec)
}

pred_ssa_sep <- function() {
  if (is.complex(signal)) {
    noise <- CCSWGN(N, 0, sd)
  } else {
    noise <- rnorm(N, 0, sd)
  }
  x <- signal + noise
  rec <- reconstruct(ssa(x, L = L_ssa), groups = groups_ssa)
  Reduce(cbind, rec)
}

pred_ssa_est <- function() {
  if (is.complex(signal)) {
    noise <- CCSWGN(N, 0, sd)
  } else {
    noise <- rnorm(N, 0, sd)
  }
  x <- signal + noise
  est <- parestimate(ssa(x, L = L_ssa), groups = groups_ssa)
  c(sort(est$freq), sort(est$rates))
}

pred_hossa_rec <- function() {
  if (is.complex(signal)) {
    noise <- CCSWGN(N, 0, sd)
  } else {
    noise <- rnorm(N, 0, sd)
  }
  x <- signal + noise
  
  rec <- tens_ssa_reconstruct(
    x,
    unlist(L_hossa),
    groups = unlist(groups_hossa, recursive = FALSE),
    svd.method = "primme",
    rm.repeated = pseudo_hosvd,
    trunc_dims = unlist(td_hossa),
    status = FALSE
  )
  
  Reduce("+", rec)
}

pred_hossa_sep <- function() {
  if (is.complex(signal)) {
    noise <- CCSWGN(N, 0, sd)
  } else {
    noise <- rnorm(N, 0, sd)
  }
  x <- signal + noise
  
  rec <- tens_ssa_reconstruct(
    x,
    unlist(L_hossa),
    groups = groups_hossa,
    svd.method = "primme",
    rm.repeated = pseudo_hosvd,
    trunc_dims = unlist(td_hossa),
    status = FALSE
  )
  
  Reduce(cbind, rec)
}

pred_nest_hossa_sep <- function() {
  if (is.complex(signal)) {
    noise <- CCSWGN(N, 0, sd)
  } else {
    noise <- rnorm(N, 0, sd)
  }
  x <- signal + noise
  neig <- max(sapply(groups_nested_hossa, max))
  
  dec <- tens_ssa_decompose(
    x,
    unlist(L_hossa),
    neig = rep(neig, 3),
    svd.method = "primme",
    rm.repeated = TRUE,
    status = FALSE
  )
  
  rec <- tens_ssa_reconstruct(dec$est,
                              groups = groups_nested_hossa,
                              svd.method = "svd",
                              status = FALSE)
  
  Reduce(cbind, rec)
}

pred_hossa_est <- function() {
  if (is.complex(signal)) {
    noise <- CCSWGN(N, 0, sd)
  } else {
    noise <- rnorm(N, 0, sd)
  }
  x <- signal + noise
  
  est <- tens_esprit(
    x,
    unlist(L_hossa),
    groups = groups_hossa,
    decomp = "HOSVD",
    svd.method = "primme",
    rm.repeated = pseudo_hosvd,
    est_dim = est_dim_hossa,
    status = FALSE
  )[[1]]
  
  c(sort(est$freq), sort(est$rates))
}

pred_mssa_rec <- function() {
  if (is.complex(signal)) {
    noise <- CCSWGN(N * P, 0, sd)
  } else {
    noise <- rnorm(N * P, 0, sd)
  }
  x <- signal + noise
  rec <- cmssa_reconstruct(x, L = L_mssa, groups = groups_mssa)
  Reduce("+", rec)
}

pred_mssa_sep <- function() {
  if (is.complex(signal)) {
    noise <- CCSWGN(N * P, 0, sd)
  } else {
    noise <- rnorm(N * P, 0, sd)
  }
  x <- signal + noise
  rec <- cmssa_reconstruct(x, L = L_mssa, groups = groups_mssa)
  Reduce(\(m1, m2) abind(m1, m2, along = 3), rec)
}

pred_mssa_est <- function() {
  if (is.complex(signal)) {
    noise <- CCSWGN(N * P, 0, sd)
  } else {
    noise <- rnorm(N * P, 0, sd)
  }
  x <- signal + noise
  est <- cmesprit(x, L = L_mssa, groups = groups_mssa)[[1]]
  c(sort(est$freq), sort(est$rates))
}

pred_homssa_rec <- function() {
  if (is.complex(signal)) {
    noise <- CCSWGN(N * P, 0, sd)
  } else {
    noise <- rnorm(N * P, 0, sd)
  }
  x <- signal + noise
  
  rec <- tens_mssa_reconstruct(
    x,
    L_homssa,
    groups = groups_homssa,
    groups3 = groups3_homssa,
    decomp = "HOSVD",
    svd.method = "primme",
    status = FALSE
  )
  
  Reduce("+", rec)
}

pred_homssa_sep <- function() {
  if (is.complex(signal)) {
    noise <- CCSWGN(N * P, 0, sd)
  } else {
    noise <- rnorm(N * P, 0, sd)
  }
  x <- signal + noise
  
  rec <- tens_mssa_reconstruct(
    x,
    L_homssa,
    groups = groups_homssa,
    groups3 = groups3_homssa,
    decomp = "HOSVD",
    svd.method = "primme",
    status = FALSE
  )
  
  Reduce(\(m1, m2) abind(m1, m2, along = 3), rec)
}

pred_homssa_est <- function() {
  if (is.complex(signal)) {
    noise <- CCSWGN(N * P, 0, sd)
  } else {
    noise <- rnorm(N * P, 0, sd)
  }
  x <- signal + noise
  
  est <- tens_esprit(
    x,
    L_homssa,
    groups = groups_homssa,
    kind = "HO-MSSA",
    decomp = "HOSVD",
    svd.method = "primme",
    rm.repeated = pseudo_hosvd,
    est_dim = est_dim_homssa,
    r3 = max.r3,
    status = FALSE
  )[[1]]
  
  c(sort(est$freq), sort(est$rates))
}
```

```{r}
calc_errors <- function(R, pred_func, err_func, true, progbar, seed = 5) {
  with(plan(multisession), {
    batch_recs <- function(R) {
      future_replicate(R, {
        progbar()
        pred_func()
      }, simplify = "array", future.seed = seed)
    }
    if (sd == 0)
      recs <- batch_recs(2) # don't need R because no noise present. Set to 2, because with 1 will be simplified
    else
      recs <- batch_recs(R)
  })
  
  return(err_func(true, recs))
}

# params_grid_list: list with the parameter names as names and vectors of the parameters values as values
calc_errors_by_params <- function(params_grid_list, R, comp_func, err_func, true, seed = 5) {
  params_grid <- expand.grid(params_grid_list)
  results <- params_grid
  results$error <- numeric(nrow(params_grid))
  q <- progressor(nrow(params_grid) * R)
  for (i in seq(nrow(params_grid))) {
    # print(paste0(names(params_grid), " = ", params_grid[i,], collapse = ", "))
    list2env(params_grid[i,], envir = parent.frame())
    results[i,]$error <- calc_errors(R, comp_func, err_func, true, q, seed)
  }
  return(results)
}
```

```{r}
print_errors <- function(err_df) {
  err_df |> mutate(zapped_error = zapsmall(error)) |> group_by(sd) |> arrange(zapped_error, L_ssa, .by_group = TRUE)
}

print_errors_m <- function(err_df) {
  err_df |> mutate(zapped_error = zapsmall(error)) |> group_by(sd) |> arrange(zapped_error, L_mssa, .by_group = TRUE)
}

summarise_errors <- function(err_df) {
  err_df |> mutate(zapped_error = zapsmall(error)) |> group_by(sd) |> summarise(
    best = min(zapped_error),
    best_L = L[which.min(zapped_error)],
    avg = mean(zapped_error),
    worst = max(zapped_error)
  ) |> arrange(sd)
}

min_errors <- function(err_df) {
  err_df |> mutate(zapped_error = zapsmall(error)) |> group_by(sd) |> 
    filter(zapped_error == min(zapped_error)) |> arrange(sd)
}
```

# 1d series reconstruction
```{r}
P_case <- ""
# S.true <- 1
S.true <- 2
S_case <- paste0("s", S.true)

# no rates
rates.true <- rep(0, S.true)
rates_case <- "_nr"
# positive small equal rates
# rates.true <- rep(0.01, S.true)
# rates_case <- "_pser"
# positive small different rates
# rates.true <- c(0.01, 0.02)
# rates_case <- "_psdr"
# small equal rated
# rates.true <- c(-0.01, -0.01)
# rates_case <- "_ser"
# large equal rates
# rates.true <- c(-0.02, -0.02)
# rates_case <- "_ler"
# different rates
# rates.true <- c(-0.01, -0.02)
# rates_case <- "_dr"

# no freqs
# freqs.true <- rep(0, S.true)
# freqs_case <- "_nf"
# same freqs
# freqs.true <- rep(0.2, S.true)
# freqs_case <- "_sf"
# freqs.true <- rep(1/12, S.true)
# freqs_case <- "_sf"
freqs.true <- c(1/12, 1/6)
freqs_case <- "_df"
# different freqs
# freqs.true <- c(0.2, 0.22)
# freqs_case <- "_df"
# close different freqs
# freqs.true <- c(0.02, 0.0205)
# freqs_case <- "_cdf"


# equal amplitudes
# A <- rep(1, S.true)
# A_case <- "_ea"
# A <- rep(30, S.true)
# A_case <- "_ea"
# different amplitudes
A <- c(2, 1)
A_case <- "_da"

poly.true <- 1 %o% rep(1, S.true)
poly_case <- ""
# poly.true <- c(1, 1) %o% rep(1, S.true)
# poly_case <- sprintf("p%d", dim(poly.true)[1] - 1)

# equal phases
phases.true <- rep(0, S.true)
phases_case <- "_ep"
# different phases
# phases.true <- c(0, pi / 3)
# phases_case <- "_dp"

# complex.signal <- TRUE
# complex_case <- "_c" # complex
complex.signal <- FALSE
complex_case <- "_r" # real

if (complex.signal) {
  true_params <- c(sort(freqs.true), sort(rates.true))
} else {
  half_idx <- which(freqs.true == 1 / 2)
  if (length(half_idx) > 0) {
    true_params <- c(sort(c(
      freqs.true[half_idx], rep(freqs.true[-half_idx], each = 2) * c(1, -1)
    )), sort(c(rates.true[half_idx], rep(rates.true[-half_idx], 2))))
  } else {
    true_params <- c(sort(rep(freqs.true, each = 2) * c(1, -1)), sort(rep(rates.true, 2)))
  }
}

# N <- 25
# n_case <- ""
# N <- 100
# N <- 71
# n_case <- "_long"
N <- 150
n_case <- "_longer"
```


## Constructing Signal
```{r}
full_case <- paste0(n_case, P_case, rates_case, freqs_case, A_case, poly_case, phases_case, complex_case)
max.r <- calc_rank(A, rates.true, freqs.true, phases.true, complex.signal, poly.true)

stopifnot(N > max.r * 2 + 1)
signal <- construct_ts(N, A, rates.true, freqs.true, phases.true, complex.signal, poly.true)
```

## Parameters grid for 1d case
```{r}
if (!is.matrix(signal)) {
groups_ssa <- list(1:max.r)

if (n_case == "") {
  Ls_ssa <- seq(from = max.r + 1,
                to = (N + 1) %/% 2,
                by = 1)
} else {
  Ls_ssa <- seq(from = max.r + 1,
                to = (N + 1) %/% 2,
                by = 5)
}

groups_hossa <- list(1:max.r)
# groups_hossa <- list(list(1:max.r), list(1:2, 3:4))
td_hossa <- 1:3
est_dim_hossa <- list(1:2, c(1, 3), 2:3, 1:3)

# I >= L >= J
if (n_case == "") {
  Is_tmp <- seq(from = max.r + 1,
                to = N - max.r - 1,
                by = 1)
  Ls_tmp <- seq(from = max.r + 1,
                to = N - max.r - 1,
                by = 1)
  Ks_tmp <- seq(from = max.r + 1,
                to = N - max.r - 1,
                by = 1)
} else {
  Is_tmp <- seq(from = max.r + 1,
                to = N - max.r - 1,
                by = 5)
  Ls_tmp <- seq(from = max.r + 1,
                to = N - max.r - 1,
                by = 5)
  Ks_tmp <- seq(from = max.r + 1,
                to = N - max.r - 1,
                by = 5)
}

correct.dims <- which(
  outer(Is_tmp, 
        Ls_tmp, 
        (\(i, l)ifelse((i >= l) & (N - i - l + 2 > max.r) & (N - i - l + 2 <= l), TRUE, FALSE))
        ), 
  arr.ind = TRUE)

len.out <- nrow(correct.dims)
tens.dims <- matrix(ncol = 3, nrow = len.out)
colnames(tens.dims) <- c("I", "L", "J")
tens.dims[, "I"] <- Is_tmp[correct.dims[, 1]]
tens.dims[, "L"] <- Ls_tmp[correct.dims[, 2]]
tens.dims[, "J"] <- N - tens.dims[, "I"] - tens.dims[, "L"] + 2

tens.dims <- tens.dims[order(tens.dims[, 1], tens.dims[, 2], decreasing = TRUE), ]
tens.dims_list <- apply(tens.dims, 1, \(v) v[1:2], simplify = list)

# TODO: Could come up with something better
tens.dims_list_4d <- list()
for (i in Is_tmp) {
  for (l in Ls_tmp) {
    if (l > i) next
    for (k in Ks_tmp) {
      j <- N - i - l - k + 3
      if (k > l || j > k || j <= max.r) next
      tens.dims_list_4d <- c(tens.dims_list_4d, list(c(i, l, k)))
    }
  }
}

# sds <- seq(0, 0.6, by = 0.2)
# sds <- c(0.6, 1.2, 1.8)
sds <- 0.6

params_grid_ssa <- list(sd = sds, L_ssa = Ls_ssa)
params_grid_hossa <- list(sd = sds,
                          L_hossa = tens.dims_list,
                          groups_hossa = groups_hossa)
params_grid_hossa_4d <- list(sd = sds,
                              L_hossa = tens.dims_list_4d,
                              groups_hossa = groups_hossa)
params_grid_hoesprit <- list(sd = sds,
                          est_dim_hossa = est_dim_hossa,
                          L_hossa = tens.dims_list)
}
```

## Calculating errors per parameters
```{r}
# R <- 500
R <- 100

ssa_err <- calc_errors_by_params(params_grid_ssa, R, pred_ssa_rec, rmse_ts_nd, signal, seed = 5)
# ssa_parest_err <- calc_errors_by_params(params_grid_ssa, R, pred_ssa_est, rmmse_params_mat, true_params, seed = 5)

# pseudo_hosvd = FALSE
# tssa_err <- calc_errors_by_params(params_grid_hossa, R, pred_hossa_rec, rmse_ts_nd, signal, seed = 5)
# tssa_parest_err <- calc_errors_by_params(params_grid_hoesprit, R, pred_hossa_est, rmmse_params_mat, true_params, seed = 5)

pseudo_hosvd = TRUE
pseudo_tssa_err <- calc_errors_by_params(params_grid_hossa, R, pred_hossa_rec, rmse_ts_nd, signal, seed = 5)

pseudo_tssa_4d_err <- calc_errors_by_params(params_grid_hossa_4d, R, pred_hossa_rec, rmse_ts_nd, signal, seed = 5)
# pseudo_tssa_parest_err <- calc_errors_by_params(params_grid_hoesprit, R, pred_hossa_est, rmmse_params_mat, true_params, seed = 5)

# save(ssa_err, ssa_parest_err, tssa_err, tssa_parest_err, pseudo_tssa_err, pseudo_tssa_parest_err, file = paste0(rprojroot::find_rstudio_root_file(), "/Research/comp_results/cases_errors/", S_case, full_case, ".RData"))
# save(ssa_err, ssa_parest_err, pseudo_tssa_err, pseudo_tssa_parest_err, file = paste0(rprojroot::find_rstudio_root_file(), "/Research/comp_results/cases_errors/", S_case, full_case, ".RData"))
save(ssa_err, pseudo_tssa_err, pseudo_tssa_4d_err, file = paste0(rprojroot::find_rstudio_root_file(), "/Research/comp_results/cases_errors/", S_case, full_case, ".RData"))
```

```{r}
min_errors(ssa_err)
min_errors(ssa_parest_err)
```

```{r}
min_errors(pseudo_tssa_err)
```

```{r}
min_errors(pseudo_tssa_4d_err)
```


## Calculating pointwise errors
```{r}
load(file = paste0(rprojroot::find_rstudio_root_file(), "/Research/comp_results/cases_errors/", S_case, full_case, ".RData"))
R <- 500
sd <- 0.6
L_ssa <- ssa_err |> filter(sd == 0.6) |> filter(error == min(error)) |> getElement("L_ssa")
td_hossa <- trunc_dims
L_hossa <- tssa_err |> filter(sd == 0.6) |> filter(error == min(error)) |> getElement("L_hossa")
ssa_err_pw <- (\() {
q <- progressor(R)
calc_errors(R, pred_ssa_rec, rmse_ts_pw, signal, q, seed = 5)
})()
tssa_err_pw <- (\() {
q <- progressor(R)
calc_errors(R, pred_hossa_rec, rmse_ts_pw, signal, q, seed = 5)
})()
save(ssa_err, tssa_err, ssa_err_pw, tssa_err_pw, file = paste0(rprojroot::find_rstudio_root_file(), "/Research/comp_results/point_wise_errors/", S_case, full_case, ".RData"))
```

```{r}
R <- 500
sd <- 0.6
L_ssa <- ssa_parest_err |> filter(sd == 0.6) |> filter(error == min(error)) |> getElement("L_ssa")

ssa_est_err_sample <- (\() {
  q <- progressor(R)
  calc_errors(R, pred_ssa_est, \(true, pred) colMeans(abs(pred - true %o% rep(1, tail(dim(pred), 1))) ^ 2), true_params, q, seed = 5)
})()

td_hossa <- trunc_dims
L_hossa <- tssa_parest_err |> filter(sd == 0.6) |> filter(error == min(error)) |> getElement("L_hossa")
est_dim_hossa <- tssa_parest_err |> filter(sd == 0.6) |> filter(error == min(error)) |> getElement("est_dim_hossa")
pseudo_hosvd = FALSE
tssa_est_err_sample <- (\() {
  q <- progressor(R)
  calc_errors(R, pred_hossa_est, \(true, pred) colMeans(abs(pred - true %o% rep(1, tail(dim(pred), 1))) ^ 2), true_params, q, seed = 5)
})()


L_hossa <- pseudo_tssa_parest_err |> filter(sd == 0.6) |> filter(error == min(error)) |> getElement("L_hossa")
L_hossa <- L_hossa[1]
est_dim_hossa <- pseudo_tssa_parest_err |> filter(sd == 0.6) |> filter(error == min(error)) |> getElement("est_dim_hossa")
est_dim_hossa <- est_dim_hossa[1]
pseudo_hosvd = TRUE
pseudo_tssa_est_err_sample <- (\() {
  q <- progressor(R)
  calc_errors(R, pred_hossa_est, \(true, pred) colMeans(abs(pred - true %o% rep(1, tail(dim(pred), 1))) ^ 2), true_params, q, seed = 5)
})()

t.test(pseudo_tssa_est_err_sample, ssa_est_err_sample, paired = TRUE, var.equal = TRUE)
```

```{r}
R <- 500
sd <- 0.6
L_ssa <- ssa_err |> filter(sd == 0.6) |> filter(error == min(error)) |> getElement("L_ssa")

ssa_err_sample <- (\() {
  q <- progressor(R)
  calc_errors(R, pred_ssa_rec, \(true, pred) colMeans(abs(pred - true %o% rep(1, tail(dim(pred), 1))) ^ 2), signal, q, seed = 5)
})()

td_hossa <- trunc_dims
L_hossa <- tssa_err |> filter(sd == 0.6) |> filter(error == min(error)) |> getElement("L_hossa")
pseudo_hosvd = FALSE
tssa_err_sample <- (\() {
  q <- progressor(R)
  calc_errors(R, pred_hossa_rec, \(true, pred) colMeans(abs(pred - true %o% rep(1, tail(dim(pred), 1))) ^ 2), signal, q, seed = 5)
})()


L_hossa <- pseudo_tssa_err |> filter(sd == 0.6) |> filter(error == min(error)) |> getElement("L_hossa")
L_hossa <- L_hossa[1]
pseudo_hosvd = TRUE
pseudo_tssa_err_sample <- (\() {
  q <- progressor(R)
  calc_errors(R, pred_hossa_rec, \(true, pred) colMeans(abs(pred - true %o% rep(1, tail(dim(pred), 1))) ^ 2), signal, q, seed = 5)
})()

t.test(pseudo_tssa_err_sample, ssa_err_sample, paired = TRUE, var.equal = TRUE)
```

## Results
```{r}
rbind(
  data.frame(rmse = ssa_err_pw, method = "SSA", n = 1:N),
  data.frame(rmse = tssa_err_pw, n = 1:N,
    method = ifelse(pseudo_hosvd, "Pseudo-HO-SSA", "HO-SSA")
    )
) |>
  ggplot() + geom_line(aes(x = n, y = rmse, color = method), linewidth = 1.4) +
  annotate("label", x = N / 4, y = max(ssa_err_pw) * 0.97,
    label = sprintf("SSA: L = %d\nHO-SSA: L = (%d, %d)", L_ssa, L_hossa[[1]][1], L_hossa[[1]][2])
  ) + theme_minimal(base_size = 15)
```

```{r}
summarise_errors(rename(ssa_err, L = L_ssa))
summarise_errors(rename(ssa_parest_err, L = L_ssa))
```

```{r}
summarise_errors(rename(tssa_err, L = L_hossa))
summarise_errors(rename(tssa_parest_err, L = L_hossa))
```

```{r}
summarise_errors(rename(pseudo_tssa_err, L = L_hossa))
summarise_errors(rename(pseudo_tssa_parest_err, L = L_hossa))
```

```{r}
# Code to check that column spaces are correct
# all(apply(as.matrix(attributes(st)$unfolds_r[[3]]), 2, \(v) any(apply(rs_unfold(st, 3)@data, 2, \(v2) all(v - v2 < 1e-8)))))
# [1] TRUE
# all(apply(rs_unfold(st, 3)@data, 2, \(v) any(apply(as.matrix(attributes(st)$unfolds_r[[3]]), 2, \(v2) all(v - v2 < 1e-8)))))
```


# MV series reconstruction
```{r}
S.true <- 2
S_case <- paste0("s", S.true)

P <- 12
P_case = "_c12"

# rates.true <- matrix(rep(0, P * S.true), nrow = P)
# rates_case <- "_nr"
# rates.true <- matrix(rep(-0.04, P * S.true), nrow = P)
# rates_case <- "_Ler"
# different by summands rates
# rates.true <- matrix(rep(-0.04, P * S.true) + rep(seq(
#   from = 0,
#   length.out = S.true,
#   by = -0.02
# ), each = P), nrow = P)
# rates_case <- "_dsr"
# different by summands positive rates
rates.true <- matrix(rep(0, P * S.true) + rep(seq(
  from = 0.01,
  length.out = S.true,
  by = 0.01
), each = P), nrow = P)
rates_case <- "_dspr"

# freqs.true <- matrix(rep(0, P * S.true), nrow = P)
# freqs_case <- "_nf"
# freqs.true <- matrix(rep(0.2, P * S.true), nrow = P)
# freqs_case <- "_ef"
freqs.true <- rep(0.2, P) %o% seq(from = 1, to = 0.22 / 0.2, length.out = S.true)
freqs_case <- "_dsf"

true_params <- c(sort(unique(freqs.true)), sort(unique(rates.true)))

# A <- matrix(rep(1, P * S.true), nrow = P)
# A_case <- "_ea"
# different by series linear amplitudes
A <- (1 + seq(1, P) / 10) %o% rep(1, S.true)
A_case <- "_dtla"
# A <- matrix(rnorm(P * S.true), nrow = P)
# A_case <- "_da"

poly.true <- 1 %o% rep(1, P) %o% rep(1, S.true)
poly_case <- ""
# poly.true <- c(1, 1) * matrix(runif(2 * P, 0.5, 2), nrow = 2) %o% rnorm(S.true, mean = 1, sd = 0.25)
# poly_case <- sprintf("p%d", dim(poly.true)[1] - 1)

phases.true <- matrix(rep(0, P * S.true), nrow = P)
phases_case <- "_ep"

complex.signal <- TRUE
complex_case <- "_c" # complex
# complex.signal <- FALSE
# complex_case <- "_r" # real

# N <- 25
# n_case <- ""
N <- 100
n_case <- "_long"
```


## Constructing Signal
```{r}
full_case <- paste0(n_case, P_case, rates_case, freqs_case, A_case, poly_case, phases_case, complex_case)
max.r <- calc_rank(A, rates.true, freqs.true, phases.true, complex.signal, poly.true)

stopifnot(N > max.r * 2 + 1)
signal <- construct_ts(N, A, rates.true, freqs.true, phases.true, complex.signal, poly.true)
```

## Parameters grid for mv case
```{r}
if (is.matrix(signal)) {
groups_mssa <- list(1:max.r)

if (n_case == "") {
  Ls_mssa <- seq(from = max.r + 1,
                to = N - max.r - 1,
                by = 2)
  Ls_homssa = seq(from = max.r + 1,
                to = (N + 1) %/% 2,
                by = 2)
} else {
  Ls_mssa <- seq(from = max.r + 1,
                to = N - max.r - 1,
                by = 10)
  Ls_homssa = seq(from = max.r + 1,
                to = (N + 1) %/% 2,
                by = 10)
}

max.r3 <- calc_rank3(signal)
groups_homssa <- list(1:max.r)
groups3_homssa <- list(1:max.r3)
est_dim_homssa <- seq(2)


# sds <- seq(0, 0.6, by = 0.2)
sds <- 0.6

params_grid_mssa <- list(sd = sds, L_mssa = Ls_mssa)
params_grid_homssa <- list(sd = sds,
                          L_homssa = Ls_homssa)
params_grid_homesprit <- list(sd = sds,
                             est_dim_homssa = est_dim_homssa,
                             L_homssa = Ls_homssa)
}
```

## Calculating errors per parameters
```{r}
R <- 500

mssa_err <- calc_errors_by_params(params_grid_mssa, R, pred_mssa_rec, rmse_ts_nd, signal, seed = 5)
mssa_parest_err <- calc_errors_by_params(params_grid_mssa, R, pred_mssa_est, rmse, true_params, seed = 5)

tmssa_err <- calc_errors_by_params(params_grid_homssa, R, pred_homssa_rec, rmse_ts_nd, signal, seed = 5)
tmssa_parest_err <- calc_errors_by_params(params_grid_homesprit, R, pred_homssa_est, rmse, true_params, seed = 5)

save(mssa_err, mssa_parest_err, tmssa_err, tmssa_parest_err, file = paste0(rprojroot::find_rstudio_root_file(), "/Research/comp_results/cases_errors/mv_", S_case, full_case, ".RData"))
```

## Result
```{r}
summarise_errors(rename(mssa_err, L = L_mssa))
summarise_errors(rename(mssa_parest_err, L = L_mssa))
```

```{r}
summarise_errors(rename(tmssa_err, L = L_homssa))
summarise_errors(rename(tmssa_parest_err, L = L_homssa))
```


# 1d series separation
```{r}
P_case <- ""
# S.true <- 1
S.true <- 2
S_case <- paste0("s", S.true)

# no rates
rates.true <- rep(0, S.true)
rates_case <- "_nr"
# positive small equal rates
# rates.true <- rep(0.01, S.true)
# rates_case <- "_pser"
# positive small different rates
# rates.true <- c(0.02, 0.01)
# rates_case <- "_psdr"
# small equal rated
# rates.true <- c(-0.01, -0.01)
# rates_case <- "_ser"
# large equal rates
# rates.true <- c(-0.02, -0.02)
# rates_case <- "_ler"
# different rates
# rates.true <- c(-0.01, -0.02)
# rates_case <- "_dr"

# no freqs
# freqs.true <- rep(0, S.true)
# freqs_case <- "_nf"
# same freqs
# freqs.true <- rep(0.2, S.true)
# freqs_case <- "_sf"
# freqs.true <- rep(1/12, S.true)
# freqs_case <- "_sf"
# different freqs
# freqs.true <- c(0.1, 0.2)
# freqs_case <- "_df"
freqs.true <- c(1/12, 1/6)
freqs_case <- "_df"
# close different freqs
# freqs.true <- c(0.02, 0.0205)
# freqs_case <- "_cdf"


# equal amplitudes
# A <- rep(1, S.true)
# A_case <- "_ea"
# A <- rep(30, S.true)
# A_case <- "_ea"
# different amplitudes
A <- c(2, 1)
A_case <- "_da"

poly.true <- 1 %o% rep(1, S.true)
poly_case <- ""
# poly.true <- c(1, 1) %o% rep(1, S.true)
# poly_case <- sprintf("p%d", dim(poly.true)[1] - 1)

# equal phases
phases.true <- rep(0, S.true)
phases_case <- "_ep"
# different phases
# phases.true <- c(0, pi / 3)
# phases_case <- "_dp"

# complex.signal <- TRUE
# complex_case <- "_c" # complex
complex.signal <- FALSE
complex_case <- "_r" # real

# N <- 25
# n_case <- ""
N <- 95
# N <- 71
n_case <- "_long"
```


## Constructing Signal
```{r}
full_case <- paste0(n_case, P_case, rates_case, freqs_case, A_case, poly_case, phases_case, complex_case)
max.r <- calc_rank(A, rates.true, freqs.true, phases.true, complex.signal, poly.true)

stopifnot(N > max.r * 2 + 1)
signal_sep <- construct_ts(N, A, rates.true, freqs.true, phases.true, complex.signal, poly.true, sep = TRUE)
signal <- apply(signal_sep, 1, sum)
```

## Parameters grid for 1d case
```{r}
if (!is.matrix(signal)) {
groups_ssa <- list(1:2, 3:4)

if (n_case == "") {
  Ls_ssa <- seq(from = max.r + 1,
                to = (N + 1) %/% 2,
                by = 1)
} else {
  Ls_ssa <- c(seq(from = max.r + 6 - (max.r + 1) %% 5,
                to = (N + 1) %/% 2,
                by = 5), 48)
}

groups_hossa <- list(1:2, 3:4)
groups_nested_hossa <- list(1:2, 3:4)
trunc_dims <- list(1:3)
est_dim_hossa <- seq(3)

# I >= L >= J
if (n_case == "") {
  Is_tmp <- seq(from = max.r + 1,
                to = N - max.r - 1,
                by = 1)
  Ls_tmp <- seq(from = max.r + 1,
                to = N - max.r - 1,
                by = 1)
} else {
  Is_tmp <- seq(from = max.r + 6 - (max.r + 1) %% 5,
                to = N - max.r - 1,
                by = 5)
  Ls_tmp <- seq(from = max.r + 6 - (max.r + 1) %% 5,
                to = N - max.r - 1,
                by = 5)
}
correct.dims <- which(outer(Is_tmp, Ls_tmp, (\(i, l) ifelse((i >= l) &
                                                      (N - i - l + 2 > max.r) &
                                                      (N - i - l + 2 <= l), TRUE, FALSE
))), arr.ind = TRUE)

len.out <- nrow(correct.dims)

tens.dims <- matrix(ncol = 3, nrow = len.out)
colnames(tens.dims) <- c("I", "L", "J")
tens.dims[, "I"] <- Is_tmp[correct.dims[, 1]]
tens.dims[, "L"] <- Ls_tmp[correct.dims[, 2]]
tens.dims[, "J"] <- N - tens.dims[, "I"] - tens.dims[, "L"] + 2

tens.dims <- tens.dims[order(tens.dims[, 1], tens.dims[, 2], decreasing = TRUE), ]
tens.dims_list <- apply(tens.dims, 1, \(v) v[1:2], simplify = list)


# sds <- seq(0, 0.6, by = 0.2)
sds <- c(0, 0.6)

params_grid_ssa <- list(sd = sds, L_ssa = Ls_ssa)
params_grid_hossa <- list(sd = sds,
                          td_hossa =  trunc_dims,
                          L_hossa = tens.dims_list)
}
```

## Calculating errors per parameters
```{r}
R <- 500

ssa_err <- calc_errors_by_params(params_grid_ssa, R, pred_ssa_sep, rmse_ts_nd, signal_sep, seed = 5)

pseudo_hosvd = FALSE
tssa_err <- calc_errors_by_params(params_grid_hossa, R, pred_hossa_sep, rmse_ts_nd, signal_sep, seed = 5)

pseudo_hosvd = TRUE
pseudo_tssa_err <- calc_errors_by_params(params_grid_hossa, R, pred_hossa_sep, rmse_ts_nd, signal_sep, seed = 5)

nested_tssa_err <- calc_errors_by_params(params_grid_hossa, R, pred_nest_hossa_sep, rmse_ts_nd, signal_sep, seed = 5)

save(ssa_err, tssa_err, pseudo_tssa_err, nested_tssa_err, file = paste0(rprojroot::find_rstudio_root_file(), "/Research/comp_results/cases_errors/sep_", S_case, full_case, ".RData"))
```

## Results
```{r}
summarise_errors(rename(ssa_err, L = L_ssa))
```

```{r}
summarise_errors(rename(tssa_err, L = L_hossa))
```

```{r}
summarise_errors(rename(pseudo_tssa_err, L = L_hossa))
```

```{r}
summarise_errors(rename(nested_tssa_err, L = L_hossa))
```